\section*{Appendix C0: No-Distance Combinatorial Austerity Line (Paper 25.1 support appendix)}

Appendix C0 collects the no-distance combinatorial branch from Stages C0.1 through C0.16 together with the first C0-to-DK bridge-validation note, without rewriting the original Phase I, Phase II, or Phase III files. It is reserved here as the support appendix for Paper 25.1. The line begins as a metric-removal control on an explicit edge-graph branch, then extends through degree-spectrum protection, incidence-noise robustness, controlled motif injection, protected sequencing, forced mismatch accumulation, kernel-inconsistency forcing, weaker bounded-path survivor maps, tri-root motif forcing, harmonic-address compatibility, resonance and detuning scans, boundary refinement, a locking-mechanism null test, a forward/reverse hysteresis probe, a selector-withdrawal retention probe, and finally a signed combinatorial bridge validation for a C0-derived Dirac--Kahler candidate. Throughout the series the graph size, representative encounter families, projected-transverse sector, and shell-weight measurement vocabulary remain fixed within each sub-branch; only the combinatorial forcing, survivor-kernel structure, harmonic-address rule, or bridge-validation operator scaffold is changed.

The combined numerical boundary is now broader and sharper. C0.6 establishes reproducible positive mismatch accumulation and protected ordering. C0.7 shows that the original adjacency versus graph-shell alternation is still too rigid to produce explicit kernel-update inconsistency. C0.8 softens that boundary: the weaker bounded-path kernel opens explicit adjacency-versus-kernel inconsistency in five of nine runs, and the hybrid noise plus motif forcing protocol produces protected ordered persistence in two of nine runs while preserving the same coarse family labels (\texttt{trapped\_local}, \texttt{smeared\_transfer}, and \texttt{split\_channel\_exchange}) across the full matrix. C0.9 then fails to scale that partial positive. The later harmonic-address branch C0.10--C0.14 opens a weak topology-selection signal on the clustered DK seed, resolves a narrow braid-to-smear detuning transition near $\delta \approx 0.75$, and closes negatively at the locking stage. C0.15 adds a partial hysteresis opening, with path dependence in four of nine runs but no broad irreversibility. C0.16 sharpens the locking null: selected topology does not survive selector withdrawal in any tested run. The bridge note then closes the signed chain identities exactly on the tested representative graph complexes and recovers a clustered-only detuning/protection signal on the derived complex, while still leaving orientation source, evolution-law equivalence, intrinsic grading, and family-wide recovery conditional.

The honest sentence to carry forward is:
\begin{quote}
\emph{The C0 line supports graph-structural class survival, reproducible incidence-driven mismatch accumulation, partial kernel inconsistency under softer survivor maps, weak harmonic-address topology selection, partial hysteresis, and a real C0-to-DK bridge scaffold, but it does not yet support broad irreversibility scaling or topology locking on the no-distance control branch.}
\end{quote}

This appendix therefore gives a clean numerical challenge to any automatic non-functoriality claim for time emergence while also separating selection from locking. The present no-distance branch supports incidence-forced protected sequencing and can be pushed into explicit inconsistency, but it does not yet support a broad claim that irreversibility follows automatically from that inconsistency on the current combinatorial control family. Nor does the later harmonic-address line support the stronger claim that selected topology locks itself into a bounded exchange regime. The bridge note extends that discipline one step further: a C0-derived graded scaffold is now executable and algebraically exact on the tested graph families, but the full graded sector is not yet derived from C0 alone.
