\documentclass[11pt]{article}

\usepackage{amsmath,amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}

\title{A Minimal Structural-Stability Oracle Based on Frozen HAOS-IIP Telemetry}
\author{Tomislav Rupi\'c \\ Independent Researcher}
\date{March 2026}

\begin{document}

\maketitle

\begin{abstract}
This paper is not a new phase. It is a tool-facing compression of the already frozen HAOS-IIP telemetry layer into one callable stability oracle. The scope is deliberately narrow: given an evolving graph-like trajectory, determine whether it remains coherent, becomes weakly deformed, or collapses into fragmentation. The public implementation is exposed as \texttt{python3 -m haos\_iip.demo stability}. It reuses only frozen telemetry for persistence, ordering, acyclicity, causal depth, and order compatibility, while using a lightweight synthetic proxy for the distance-coherence readout in the public example layer. On the default deterministic bundle, the classifier separates three synthetic scenarios into the intended ladder \texttt{baseline $\to$ stable}, \texttt{perturbed $\to$ marginal}, and \texttt{fragmented $\to$ unstable}. On a small deterministic scan over noise and connectivity-drop parameters, the baseline scenario yields $3$ stable cells, $0$ marginal cells, and $6$ unstable cells. The bounded conclusion is operational only: frozen HAOS-IIP telemetry can act as a structural-stability oracle for small evolving graph-like systems. No claim is made about continuum limits, spacetime emergence, metric geometry, or ontology.
\end{abstract}

\section{Problem}

The immediate problem addressed here is practical rather than ontological. The question is not whether a new phase should be opened or whether deeper emergence claims are justified. The question is:
\[
\text{given a trajectory, can the frozen telemetry layer decide whether the structure remains stable?}
\]

That question matters because it turns the repository from a research archive into a callable diagnostic surface. A tool that distinguishes coherent, weakly deformed, and fragmented regimes is already useful even if no additional physics program is pursued.

\section{Frozen Telemetry Basis}

The oracle is built only on the already frozen telemetry layer in:
\begin{itemize}
\item \texttt{telemetry/frozen\_metrics.py}.
\end{itemize}

The core frozen computations reused directly are:
\begin{itemize}
\item persistence time;
\item event-threshold arrival ordering;
\item reconstructed influence edges;
\item acyclicity score;
\item causal-depth extraction;
\item order compatibility.
\end{itemize}

For public readability, the user-facing output applies one alias layer:
\begin{itemize}
\item persistence $\mapsto$ structural retention;
\item ordering $\mapsto$ temporal consistency;
\item depth drift $\mapsto$ causal deformation;
\item distance coherence $\mapsto$ geometric integrity.
\end{itemize}

The first three remain direct frozen telemetry outputs. The fourth is a lightweight synthetic proxy used only in the demo layer so a small public ladder can be shown without touching any frozen phase bundle.

\section{Synthetic Scenario Surface}

The public demo consumes three deterministic scenarios:
\begin{itemize}
\item \texttt{baseline};
\item \texttt{perturbed};
\item \texttt{fragmented}.
\end{itemize}

These are stored in:
\begin{itemize}
\item \texttt{examples/scenarios/baseline.json};
\item \texttt{examples/scenarios/perturbed.json};
\item \texttt{examples/scenarios/fragmented.json}.
\end{itemize}

The demo also exposes deterministic generator controls:
\begin{itemize}
\item \texttt{--noise};
\item \texttt{--connectivity-drop};
\item \texttt{--cluster-split}.
\end{itemize}

These modifiers perturb only the synthetic scenario layer. They do not edit, reinterpret, or reopen any frozen authority phase.

\section{CLI Interface}

The primary public entrypoint is:
\begin{verbatim}
python3 -m haos_iip.demo stability
\end{verbatim}

Machine-readable mode is available through:
\begin{verbatim}
python3 -m haos_iip.demo stability baseline --json
\end{verbatim}

The JSON payload includes:
\begin{itemize}
\item the scenario name;
\item the classification label;
\item structural retention;
\item temporal consistency;
\item causal deformation;
\item geometric integrity;
\item the underlying technical metrics;
\item artifact paths for the generated CSV and SVG outputs.
\end{itemize}

Deterministic parameter-space exploration is available through scan mode:
\begin{verbatim}
python3 -m haos_iip.demo stability \
  --scan noise=0.00:0.10:0.05 connectivity=0.00:0.20:0.10
\end{verbatim}

This writes a scan-grid CSV and a heatmap SVG while preserving the same frozen telemetry basis.

\section{Classification Ladder}

The default deterministic bundle yields the following compact result:

\begin{center}
\begin{tabular}{lccccr}
\toprule
scenario & retention & consistency & deformation & integrity & class \\
\midrule
baseline   & 1.000 & 1.000 & 0.000 & 1.000 & stable \\
perturbed  & 0.750 & 0.833 & 0.062 & 0.625 & marginal \\
fragmented & 0.250 & 0.656 & 0.172 & 0.375 & unstable \\
\bottomrule
\end{tabular}
\end{center}

This is the smallest useful result. The classifier does not just print raw telemetry. It organizes the outputs into an interpretable ladder that a human or an agent can consume quickly.

\section{Deterministic Scan Behavior}

The small scan
\[
\texttt{noise} \in \{0.00,0.05,0.10\},
\qquad
\texttt{connectivity\_drop} \in \{0.00,0.10,0.20\}
\]
was evaluated on the \texttt{baseline} scenario.

The resulting classification counts are:
\begin{itemize}
\item stable: $3$;
\item marginal: $0$;
\item unstable: $6$.
\end{itemize}

That scan is intentionally tiny, but it already changes the character of the tool. The system is no longer a static example only; it becomes a deterministic parameter-space explorer over a small synthetic regime.

\section{Why This Matters}

The oracle demonstrates three practical properties at once:
\begin{itemize}
\item frozen telemetry can be reused without opening new phases;
\item the outputs are interpretable enough for a human-facing command-line tool;
\item the interface is structured enough for downstream agent or pipeline use.
\end{itemize}

In that sense the result is not ``new physics.'' It is the first usable unit extracted from the broader HAOS-IIP program.

\section{Limits and Non-Claims}

This oracle should be interpreted as:
\begin{itemize}
\item a structural-stability diagnostic;
\item a callable telemetry surface;
\item a deterministic synthetic scenario explorer.
\end{itemize}

It should not be interpreted as:
\begin{itemize}
\item a new frozen authority phase;
\item evidence of continuum limits;
\item evidence of metric geometry;
\item evidence of spacetime emergence;
\item evidence of physical ontology.
\end{itemize}

The public distance readout in particular is a demo-layer proxy, not a new geometrical claim.

\section{Conclusion}

Frozen HAOS-IIP telemetry can be compressed into a minimal structural-stability oracle. The resulting tool distinguishes coherent, weakly perturbed, and strongly fragmented synthetic dynamics; exposes human-readable and machine-readable outputs; and supports a small deterministic scan surface without modifying any frozen phase bundle. That is sufficient to establish a practical diagnostic capability. It is not sufficient to establish ontology.

\end{document}
